### 2018-02-23 thesis/contents/tn_sagit.tex
Updates from ShareLaTeX

The trigeminal sensory system involves both the discriminative touch and pain pathways. Anatomically, the discriminative pathway synapses onto the primary trigeminal sensory nucleus, and then projects to the contralateral VC thalamus and then to the S1 cortex. The nociceptive pathway synapses caudally onto the trigeminal spinal nucleus, decussates at the level of the medulla and spinal cord. The pathway merges with the discriminative sensory pathway and ascend as the contralateral trigeminal lemniscus, and also project to the contralateral VC thalamus. The pain pathway also projects to the medial nucleus of the thalamus, where it then project medially to the cingulate cortex.

@@

\subsubsection{Along-the-tract Analysis}
Tractography centroids, which were the average curve best fit the tract bundle, were identified on the merged tracts in template space. The spline was then resampled into 30 points and became the centroid definition. Nearest-neighbor point assignment was used to divide the merged tract into 30 sections along the centroid. Diffusion metrics from the tractography points were averaged for each of the sections. CN V centroids were manually defined using 3 control points. A Catmull-Rom spline \cite{DeRose1988} was then defined from these control points using the VTK software library \cite{Schroeder2005}. TPT and S1 centroids were defined using QuickBundles \cite{Garyfallidis2012} with distance threshold of 30.

@@

\subsubsection{Diffusivity Differences}
\paragraph{CN V}
The classifier distinguished bilateral trigeminal nerves (affected and unaffected) from controls as well as affected trigeminal nerve from unaffected nerves.Both sides were distinguished from controls with classifiers that achieved around 80\% accuracy (Figure \ref{fig:GPfigure4}). 
Classifiers on the right side had accuracie ranging from 76\% (GFA) to 86\% (FA). The left classifiers achieved accuracy ranging from 79\% (MD) to 81\% (FA). The right side showed more overlap of segment ranges between positions 14--24, in the region of the cistern/gangion; while the left side showed segments that covered the entire the nerve, and extended to cover position 0--4 in the brainstem. Overall, the right side classifiers extend further into the distal space, into the cistern and even the ganglion, while the left side extended more into the brainstem. FA classifier consistently achieved the highest accuracy in both sides, with right side at position 17--23 (cistern/ganglion) and left FA at 2--16 (brainstem/cister).
The FA classifier achieved 75\% accuracy in differentiating affected versus unaffected nerves, where its segments ranged from 11 to 19, which was the cisternal segment. The GFA classifier achieved 70\% accuracy in region 9--12. Demonstrating that affected-unaffected diffusivity differences were found in the cistern/REZ (Table \ref{table:CN V}).

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure-GP-CNV.pdf}
\caption{CN V classifier performance}
\label{fig:GPfigure4}
\end{figure}

@@

\paragraph{Trigeminopontothalamic tract}
Both left and right TPT pathways could be distinguished from controls with greater than 80\% accuracy (Figure \ref{fig:GPfigureTPT}). The right maximum accuracy was 85\% (FA), and minimum accuracy 83\% (RD), while the left maximum accuracy was 89\% (FA) and minimum 71\% (RD). The regions covered both the decussation and the ascending segments of the TPT, with the right decussation showing more proximal coverage. The left-right classifiers achieved maximum accuracy of 98\% (FA) and minimum of 88\% (MD). The left-right FA covered position 4--14, suggesting strong laterality differences in the TPT (Table \ref{table:tpt}).

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure-GP-TPT.pdf}
\caption{TPT classifier performance}
\label{fig:GPfigureTPT}
\end{figure}

\begin{table}[ht]
\centering
\csvautotabular{tpt_table.csv}
\caption{TPT GP classifiers performance data}
\caption*{List of the best accuracies for each diffusion metric. Precision, recall, and f1 scores are also provided for reference}
\label{table:tpt}
\end{table}

\paragraph{S1}
Both sides of the S1 face region pathway were able to be distinguished by classifiers with about 85\% accuracy. The left RD classifier achieved max accuracy of 87\% at position 13--20, while the right RD classifier achieved max accuracy of 83\% at psotion 8--26 (Figure \ref{fig:GPfigure5}). Overall the left classifier ranged from 72\% (AD) to 87\% (RD), and the right classifiers ranged from 73\% (GFA) to 83\% (RD) (Table \ref{table:s1}). Both sides differ from controls in the branching region between the thalamus and cortex. Left-right classifiers showed maximum accuracy of 78\% (RD), and minimum of 76\% (FA). The left-right classifiers overlap in the regions 5--10, at the segment where the pathway emerges from the thalamus.

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure-GP-S1.pdf}
\caption{S1 classifier performance}
\label{fig:GPfigure5}
\end{figure}

>>>>

The trigeminal sensory system involves both the discriminative touch and pain pathways.  Pain and discriminitive touch stimuli trigger innervations of the three trigeminal nerve branches (V1, V2, and V3) in the head and face, these primary affrents converge and form the trigeminal root gangion (TRG) (also known as Gasserian, or semilunar ganglion) in the middle cranial fossa. The discriminitve touch afferents then pass through the cisternal segment of CN V, which then enters the pons, and synapse onto the primary trigeminal sensory nucleus, and then projects to the contralateral VC thalamus and then to the S1 cortex. The nociceptive pathway synapses caudally onto the trigeminal spinal nucleus, decussates at the level of the medulla and spinal cord. The pathway merges with the discriminative sensory pathway and ascend as the contralateral trigeminal lemniscus, and also project to the contralateral VC thalamus. The pain pathway also projects to the medial nucleus of the thalamus, where it then project medially to the cingulate cortex.

@@

\subsubsection{Along-the-tract Analysis}
For the purpose of quantifying the trigeminal pathways, a centroid was defined for the merged tractography of each pathway (Figure \ref{fig:GPfigure-method-centroid}). The centroids were defined as the average curve that best fit the tract bundles, and were in template space. CN V centroids were manually defined using 3 control points. A Catmull-Rom spline \cite{DeRose1988} was then defined from these control points using the VTK software library \cite{Schroeder2005}. The spline was then resampled into 30 points. TPT and S1 centroids were defined using QuickBundles \cite{Garyfallidis2012} with distance threshold of 30. Nearest-neighbor point assignment was used to cluster  the points of the merged tract model to the centroid control points, to obtain the corresponding 30 sections along each centroid. Diffusion metrics from the tractography points were averaged for each of the sections for each subject.

@@

\subsubsection{Diffusivity Differences}
\paragraph{CN V}
The classifier distinguished bilateral trigeminal nerves (affected and unaffected) from controls as well as affected trigeminal nerve from unaffected nerves.Both sides were distinguished from controls with classifiers that achieved around 80\% accuracy (Figure \ref{fig:GPfigure4}). 
Classifiers on the right side had accuracies ranging from 76\% (GFA) to 86\% (FA). The left classifiers achieved accuracy ranging from 79\% (MD) to 81\% (FA). The right side showed more overlap of segment ranges between positions 14--24, in the region of the cistern/ganglion; while the left side showed segments that covered the entire the nerve, and extended to cover position 0--4 in the brainstem. Overall, the right side classifiers extend further into the distal space, into the cistern and even the ganglion, while the left side extended more into the brainstem. FA classifier consistently achieved the highest accuracy in both sides, with right side at position 17--23 (cistern/ganglion) and left FA at 2--16 (brainstem/cistern).
The FA classifier achieved 75\% accuracy in differentiating affected versus unaffected nerves, where its segments ranged from 11 to 19, which was the cisternal segment. The GFA classifier achieved 70\% accuracy in region 9--12. Demonstrating that affected-unaffected diffusivity differences were found in the cistern/REZ (Table \ref{table:CN V}).

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure-GP-CNV.pdf}
\caption{CN V classifier performances. The left panel shows the coverage (y-axis) and accuracies (x-axis) of the GP classifiers with greater than 70\% accuracy. The middle panel is a heatmap of the FA values of the corresponding centroid positions. The right panel shows the centroid and point clusterings for anatomical reference. The colours of the points correspond to their respective cluster index on a rainbow spectrum. }
\label{fig:GPfigure4}
\end{figure}

@@

\paragraph{Trigeminopontothalamic tract}
Classifiers could distinguish bilateral TPT pathways (affected and unaffected) from controls with greater than 80\% accuracy (Figure \ref{fig:GPfigureTPT}). The right FA classifier achieved the greatest accuracy at 85\%, and right RD classifier was least accuracy at 83\%. The left FA classifier showed greatest accuracy at 89\% and left RD classifier showed the least accuray at 71\%. The classifiers covered both the decussation and the ascending segments of the TPT, with the right decussation showing more proximal coverage. The left-versus-right classifiers achieved maximum accuracy of 98\% (FA) and minimum of 88\% (MD). The FA classifier covered position 4--14, suggesting strong laterality differences in the TPT (Table \ref{table:tpt}).

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure-GP-TPT.pdf}
\caption{TPT classifier performance, see Figure \ref{fig:GPfigure4} for details on annotations.}
\label{fig:GPfigureTPT}
\end{figure}

\begin{table}[ht]
\centering
\csvautotabular{tpt_table.csv}
\caption*{List of the best accuracies for each diffusion metric. Precision, recall, and f1 scores are also provided for reference}
\label{table:tpt}
\end{table}

\paragraph{S1}
Classifiers distinguished bilateral S1 face region pathways from controls with about 85\% accuracy. The left RD classifier achieved max accuracy of 87\% at position 13--20, while the right RD classifier achieved max accuracy of 83\% at position 8--26 (Figure \ref{fig:GPfigure5}). Overall the left classifier ranged from 72\% (AD) to 87\% (RD), and the right classifiers ranged from 73\% (GFA) to 83\% (RD) (Table \ref{table:s1}). Classifiers of both sides covered the the branching region between the thalamus and cortex. Left-versus-right classifiers showed maximum accuracy of 78\% (RD), and minimum of 76\% (FA). The left-versus-right classifiers overlap in the regions 5--10, at the segment where the pathway emerges from the thalamus.

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure-GP-S1.pdf}
\caption{S1 classifier performance, see Figure \ref{fig:GPfigure4} for details on annotations.}
\label{fig:GPfigure5}
\end{figure}
